A template for the \emph{arxiv} style

### 2021-10-12 main.Rmd
more data structure section, new create a cubble section, files to top level

Spatio-temporal data can be thought of as characterised by three dimensions: group, time, and variable, which allows us to use a cube to represent the data. While group and time are can be seen as identifiers of the record, variables can be further divided into those that are group-related and time-related. Group-related variables are invariant to the time and have only one value per group while time-related variables change in time and requires both group and time to identify a record. Hence the cubic representation can be further simplified for those group-related variables as in the sketch \ref{fig:framework}, where group-related variables are squashed in the time dimension.

```{r framework, echo = FALSE, fig.cap = "Cubic representation of spatio-temporal data. (1) shows the full cube, (2) simplifies teh cube by squashing the time dimension of time-invariant variables. With (2), a slice along the group dimension will affect both time-variant and invariant variables as in (3) while, a slice along the time will only affect the tim-variant variables as in (4)."}
knitr::include_graphics(here::here("figures/cubble-framework-hand.jpeg"))
```

A cubble simplifies the workflow in spatio-temporal data through managing group and time-related variables separately in two forms: list-column and long form. The nested form:

@@

This form focuses on the time-invariant variable by squashing the time dimension

In the long form,

  - each combination of group and timestamp occupies a row
  - time-related variables are displayed, and 
  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute

Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020. Notice that each station forms a group in both forms and specifically, the nested  and long form have a underlying `rowwise_df` and `grouped_df` respectively.

```{r echo = FALSE}
climate_small
```

```{r echo = FALSE}
climate_small %>% stretch()
```

With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way. The table below shows how some `dplyr` verbs are mapped into the operation in a cubble. With the existing grouping on the station, additional groupping can be added with `group_by` and removed with `ungrouped`. [talk about why it is useful]

```{r verbs, echo = FALSE, out.height = "50%", out.width = "100%", fig.cap = "sdfsdf"}
knitr::include_graphics(here::here("figures/cubble-verbs.png"))
```

## Cubble verbs

Mention different types of manipulation with cubble:

  - `dplyr` support for cubble: 
    - basic 5s: mutate, filter, summarise, select, arrange
    - group and ungroup: group_by, ungroup
    - slice family
  - summarise missing stats

@@

```{r}
climate_small %>% 
  stretch() %>% 
  filter(year(date) == 2020) %>% 
  tamp() %>% 
  mutate(tmax_missing = ifelse(all(is.na(ts$tmax)), TRUE, FALSE)) %>% 
  filter(!tmax_missing)
```

@@

```{r}
# join wmo_id for each station
to_join <- station %>% select(id, wmo_id)
out <- climate_small %>% 
  left_join(to_join, by = c("station" = "id")) %>% 
  filter(!is.na(wmo_id))
out
```

@@

```{r}
out %>% 
  stretch() %>% 
  migrate(station, lat, long)
```

>>>>

Spatio-temporal data don't usually come to the analysts as a whole piece. A way to look at these data is to divide it into spatial and temporal dimension with an ID that links between the two. The first row in Figure \ref{fig:cubble-diagram} illustrates this representation where in the spatial dimension, the data is characterised by `id`, `lat`, `long`. $V_s$ in the last column represents all the other site-wise variables, for example, elevation and full name etc. The temporal dimension, on the other hand, can be characterised by `id` and `t` with $V_t$ representing all the time-wise variables. In climate data, this could include precipitation, maximum or minimum temperature, and wind speed etc.

```{r cubble-diagram, echo = FALSE, fig.cap="Cubble diagram", out.height="40%", fig.align='center'}
knitr::include_graphics(here::here("figures/cubble-diagram/cubble-diagram.001.png"))
```

To work with spatio-temporal data, analysts can choose to either work separately on each dimension or join the two sets together, however, each approach has its own problem: While is is natural to work separately on each sheet (since spatial and temporal operations usually don't overlap), analysts will need to manually keep the other data frame up to date. For example, the following pseudo code illustrates the scenario where once the spatial dataset is filtered for those within Victoria, the temporal dataset needs to be manually updated to reflect this spatial filter.

```{r eval = FALSE}
spatial_new <- spatial %>% filter(SITES_IN_VICTORIA)
temporal_new <- temporal %>% filter(id %in% spatial_new$id)
```

If analysts choose to join the spatial and temporal data together, the joined dataset could be too large since each spatial variable will be repeated at each time stamp for each site. Also, recordings of the site ID from different data sources can be slightly different from each other, causing a painful checking and cleaning of site IDs before the join.

A cubble, in essence, wires both dimensions in the spatio-temporal data into one object while provide two forms for manipulation the spatial and temporal dimension separately.

When manipulating the spatial dimension it uses a nest form that:

@@

When manipulating the temporal dimension, it uses the long form that:

  - each combination of group and timestamp occupies a row
  - time-related variables are displayed, and 
  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute.

# Create a cubble

The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier. `climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier. To create a cubble for this data, use:

```{r}
climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
```

Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata. For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list. As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne. We first create a station data farme with the 5 target stations.

```{r echo = FALSE}
latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631)

cand <- aus_climate %>% 
  select(id, lat, long, elev, name, wmo_id) %>% 
   as_tibble()

calc_dist <- function(df_ref, df, n = 5){
  long_ref <- df_ref$long
  lat_ref <- df_ref$lat

  df %>% 
    mutate(long_ref = long_ref,
           lat_ref=  lat_ref,
           dist = rnoaa::meteo_spherical_distance(lat, long,lat_ref, long_ref)) %>% 
    slice_min(dist, n = n) %>% 
    mutate(city = df_ref$id) %>% 
    select(-long_ref, -lat_ref)
}

(stations <- do.call("rbind", purrr::map(1:nrow(latlon_ref), 
                             ~calc_dist(latlon_ref[.x, ], cand))))
```

We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example.

```{r eval = FALSE}
sydmel_climate <- stations %>% 
  rowwise() %>% 
  mutate(ts = list(meteo_pull_monitors(id, 
                                       date_min = "2020-01-01", 
                                       date_max = "2020-12-31",
                                       var = c("PRCP", "TMAX", "TMIN")) %>% 
                     select(-id))) %>% 
  as_cubble(key = id, index = date, coords = c(long, lat))
```

```{r echo = FALSE}
(sydmel_climate <- stations %>% 
   left_join(aus_climate) %>% 
   as_cubble(key = id, index = date, coords = c(long, lat)))
```

<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020. Notice that each station forms a group in both forms and specifically, the nested  and long form have a underlying `rowwise_df` and `grouped_df` respectively. -->

<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way. The table below shows how some `dplyr` verbs are mapped into the operation in a cubble. With the existing grouping on the station, additional groupping can be added with `group_by` and removed with `ungrouped`. [talk about why it is useful] -->

## Cubble operations

### Basics

- `stretch`: nest to long form
- `tamp`: long to nest form
- `migrate`: move selected spatial variables to the long form.
- `add_dscrb_prct`: summary stats for missingness

dplyr compatibility:

  - mutate, filter, summarise, select, arrange
  - group and ungroup: group_by, ungroup
  - slice family

### Combine two cubbles

  - match river and weather gauges data
  - involve combining two cubbles
  - join operations combine the two together by appending more rows but what we really want is to bind rows.
  - bind rows also doesn't work since we want to bind only when there' s a matching????
  - introduce bind_join

### Hierarchical structure in cubble

  - hierarchical is common. 
  - Given examples. 
  - Essence: switch between different levels
  - introduce `switch_key`

@@

```{r}
aus_climate %>% 
  stretch() %>% 
  filter(year(date) == 2020) %>% 
  tamp()
```

@@

```{r}
# join wmo_id for each station
# to_join <- station %>% select(id, wmo_id)
# out <- climate_small %>% 
#   left_join(to_join, by = c("station" = "id")) %>% 
#   filter(!is.na(wmo_id))
# out
```

@@

```{r}
aus_climate %>% 
  stretch() %>% 
  migrate(id, lat, long)
```

### 2021-10-12 main.tex
more data structure section, new create a cubble section, files to top level

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

@@

Spatio-temporal data can be thought of as characterised by three
dimensions: group, time, and variable, which allows us to use a cube to
represent the data. While group and time are can be seen as identifiers
of the record, variables can be further divided into those that are
group-related and time-related. Group-related variables are invariant to
the time and have only one value per group while time-related variables
change in time and requires both group and time to identify a record.
Hence the cubic representation can be further simplified for those
group-related variables as in the sketch \ref{fig:framework}, where
group-related variables are squashed in the time dimension.

\begin{figure}
\includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/PhD/research/paper-cubble/figures/cubble-framework-hand} \caption{Cubic representation of spatio-temporal data. (1) shows the full cube, (2) simplifies teh cube by squashing the time dimension of time-invariant variables. With (2), a slice along the group dimension will affect both time-variant and invariant variables as in (3) while, a slice along the time will only affect the tim-variant variables as in (4).}\label{fig:framework}
\end{figure}

A cubble simplifies the workflow in spatio-temporal data through
managing group and time-related variables separately in two forms:
list-column and long form. The nested form:

@@

This form focuses on the time-invariant variable by squashing the time
dimension

In the long form,

\begin{itemize}
\tightlist
\item
  each combination of group and timestamp occupies a row
\item
  time-related variables are displayed, and
\item
  group-related variables are not explicitly displayed but can be
  accessed through the \texttt{meta} attribute
\end{itemize}

Below are the how the nested and long form look like for Australia
climate data, which records daily precipitation, maximum and minimum
temperature for 55 stations across Australia from 2015- 2020. Notice
that each station forms a group in both forms and specifically, the
nested and long form have a underlying \texttt{rowwise\_df} and
\texttt{grouped\_df} respectively.

\begin{verbatim}
## # Cubble: station-wise: nested form
## # Group:  station [55]
##    station       lat  long elevation name                ts                  
##    <fct>       <dbl> <dbl>     <dbl> <fct>               <list>              
##  1 ASN00001019 -14.3  127.      23   kalumburu           <tbl_ts [2,147 x 4]>
##  2 ASN00002012 -18.2  128.     422   halls creek airport <tbl_ts [2,191 x 4]>
##  3 ASN00003003 -17.9  122.       7.4 broome airport      <tbl_ts [2,192 x 4]>
##  4 ASN00006011 -24.9  114.       4   carnarvon airport   <tbl_ts [2,192 x 4]>
##  5 ASN00009021 -31.9  116.      15.4 perth airport       <tbl_ts [2,192 x 4]>
##  6 ASN00009193 -32.0  116.      43.1 rottnest island     <tbl_ts [2,192 x 4]>
##  7 ASN00009518 -34.4  115.      13   cape leeuwin        <tbl_ts [2,184 x 4]>
##  8 ASN00009789 -33.8  122.      25   esperance           <tbl_ts [2,187 x 4]>
##  9 ASN00010286 -31.6  117.     217.  cunderdin airfield  <tbl_ts [2,190 x 4]>
## 10 ASN00010917 -32.7  117.     275   wandering           <tbl_ts [2,192 x 4]>
## # ... with 45 more rows
\end{verbatim}

\begin{verbatim}
## # Cubble: time-wise: long form [tsibble]
## # Group:  station [55]
##    station     date        prcp  tmax  tmin
##    <fct>       <date>     <dbl> <dbl> <dbl>
##  1 ASN00001019 2015-01-01   164  31.5  25  
##  2 ASN00001019 2015-01-02   124  31.3  25  
##  3 ASN00001019 2015-01-03    50  29.2  24.7
##  4 ASN00001019 2015-01-04   204  32.1  24.3
##  5 ASN00001019 2015-01-05   412  31.4  24.6
##  6 ASN00001019 2015-01-06   200  28.7  25.1
##  7 ASN00001019 2015-01-07   822  30.2  24.5
##  8 ASN00001019 2015-01-08    22  31    26.6
##  9 ASN00001019 2015-01-09    62  32.7  26.2
## 10 ASN00001019 2015-01-10   274  32.9  22.6
## # ... with 120,311 more rows
\end{verbatim}

With a cubic framework on mind, different types of manipulation with
cubble can be thought of as slicing the cube in various way. The table
below shows how some \texttt{dplyr} verbs are mapped into the operation
in a cubble. With the existing grouping on the station, additional
groupping can be added with \texttt{group\_by} and removed with
\texttt{ungrouped}. {[}talk about why it is useful{]}

\begin{figure}
\includegraphics[width=1\linewidth,height=0.5\textheight]{/Users/hzha400/Documents/PhD/research/paper-cubble/figures/cubble-verbs} \caption{sdfsdf}\label{fig:verbs}
\end{figure}

\hypertarget{cubble-verbs}{%
\subsection{Cubble verbs}\label{cubble-verbs}}

Mention different types of manipulation with cubble:

\begin{itemize}
\tightlist
\item
  \texttt{dplyr} support for cubble:

  \begin{itemize}
  \tightlist
  \item
    basic 5s: mutate, filter, summarise, select, arrange
  \item
    group and ungroup: group\_by, ungroup
  \item
    slice family
  \end{itemize}
\item
  summarise missing stats
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{climate\_small }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{==} \DecValTok{2020}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{tamp}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{tmax\_missing =} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{all}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{tmax)), }\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{tmax\_missing)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: station-wise: nested form
## # Group:  station [55]
##    station      lat  long elevation name             ts             tmax_missing
##    <fct>      <dbl> <dbl>     <dbl> <fct>            <list>         <lgl>       
##  1 ASN000010~ -14.3  127.      23   kalumburu        <tbl_ts [366 ~ FALSE       
##  2 ASN000020~ -18.2  128.     422   halls creek air~ <tbl_ts [366 ~ FALSE       
##  3 ASN000030~ -17.9  122.       7.4 broome airport   <tbl_ts [366 ~ FALSE       
##  4 ASN000060~ -24.9  114.       4   carnarvon airpo~ <tbl_ts [366 ~ FALSE       
##  5 ASN000090~ -31.9  116.      15.4 perth airport    <tbl_ts [366 ~ FALSE       
##  6 ASN000091~ -32.0  116.      43.1 rottnest island  <tbl_ts [366 ~ FALSE       
##  7 ASN000095~ -34.4  115.      13   cape leeuwin     <tbl_ts [366 ~ FALSE       
##  8 ASN000097~ -33.8  122.      25   esperance        <tbl_ts [361 ~ FALSE       
##  9 ASN000102~ -31.6  117.     217.  cunderdin airfi~ <tbl_ts [366 ~ FALSE       
## 10 ASN000109~ -32.7  117.     275   wandering        <tbl_ts [366 ~ FALSE       
## # ... with 45 more rows
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# join wmo\_id for each station}
\NormalTok{to\_join }\OtherTok{\textless{}{-}}\NormalTok{ station }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(id, wmo\_id)}
\NormalTok{out }\OtherTok{\textless{}{-}}\NormalTok{ climate\_small }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{left\_join}\NormalTok{(to\_join, }\AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"station"} \OtherTok{=} \StringTok{"id"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wmo\_id))}
\NormalTok{out}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: station-wise: nested form
## # Group:  station [51]
##    station       lat  long elevation name              ts                 wmo_id
##    <fct>       <dbl> <dbl>     <dbl> <fct>             <list>              <dbl>
##  1 ASN00001019 -14.3  127.      23   kalumburu         <tbl_ts [2,147 x ~  94100
##  2 ASN00002012 -18.2  128.     422   halls creek airp~ <tbl_ts [2,191 x ~  94212
##  3 ASN00003003 -17.9  122.       7.4 broome airport    <tbl_ts [2,192 x ~  94203
##  4 ASN00006011 -24.9  114.       4   carnarvon airport <tbl_ts [2,192 x ~  94300
##  5 ASN00009021 -31.9  116.      15.4 perth airport     <tbl_ts [2,192 x ~  94610
##  6 ASN00009193 -32.0  116.      43.1 rottnest island   <tbl_ts [2,192 x ~  94602
##  7 ASN00009518 -34.4  115.      13   cape leeuwin      <tbl_ts [2,184 x ~  94601
##  8 ASN00009789 -33.8  122.      25   esperance         <tbl_ts [2,187 x ~  94638
##  9 ASN00010286 -31.6  117.     217.  cunderdin airfie~ <tbl_ts [2,190 x ~  95625
## 10 ASN00010917 -32.7  117.     275   wandering         <tbl_ts [2,192 x ~  95640
## # ... with 41 more rows
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{migrate}\NormalTok{(station, lat, long)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: time-wise: long form
## # Group:  station [55]
##    station     date        prcp  tmax  tmin   lat  long
##    <fct>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
##  1 ASN00001019 2015-01-01   164  31.5  25   -14.3  127.
##  2 ASN00001019 2015-01-02   124  31.3  25   -14.3  127.
##  3 ASN00001019 2015-01-03    50  29.2  24.7 -14.3  127.
##  4 ASN00001019 2015-01-04   204  32.1  24.3 -14.3  127.
##  5 ASN00001019 2015-01-05   412  31.4  24.6 -14.3  127.
##  6 ASN00001019 2015-01-06   200  28.7  25.1 -14.3  127.
##  7 ASN00001019 2015-01-07   822  30.2  24.5 -14.3  127.
##  8 ASN00001019 2015-01-08    22  31    26.6 -14.3  127.
##  9 ASN00001019 2015-01-09    62  32.7  26.2 -14.3  127.
## 10 ASN00001019 2015-01-10   274  32.9  22.6 -14.3  127.
## # ... with 111,562 more rows
\end{verbatim}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

@@

Spatio-temporal data don't usually come to the analysts as a whole
piece. A way to look at these data is to divide it into spatial and
temporal dimension with an ID that links between the two. The first row
in Figure \ref{fig:cubble-diagram} illustrates this representation where
in the spatial dimension, the data is characterised by \texttt{id},
\texttt{lat}, \texttt{long}. \(V_s\) in the last column represents all
the other site-wise variables, for example, elevation and full name etc.
The temporal dimension, on the other hand, can be characterised by
\texttt{id} and \texttt{t} with \(V_t\) representing all the time-wise
variables. In climate data, this could include precipitation, maximum or
minimum temperature, and wind speed etc.

\begin{figure}

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-diagram/cubble-diagram.001}

}

\caption{Cubble diagram}\label{fig:cubble-diagram}
\end{figure}

To work with spatio-temporal data, analysts can choose to either work
separately on each dimension or join the two sets together, however,
each approach has its own problem: While is is natural to work
separately on each sheet (since spatial and temporal operations usually
don't overlap), analysts will need to manually keep the other data frame
up to date. For example, the following pseudo code illustrates the
scenario where once the spatial dataset is filtered for those within
Victoria, the temporal dataset needs to be manually updated to reflect
this spatial filter.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{spatial\_new }\OtherTok{\textless{}{-}}\NormalTok{ spatial }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(SITES\_IN\_VICTORIA)}
\NormalTok{temporal\_new }\OtherTok{\textless{}{-}}\NormalTok{ temporal }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{\%in\%}\NormalTok{ spatial\_new}\SpecialCharTok{$}\NormalTok{id)}
\end{Highlighting}
\end{Shaded}

If analysts choose to join the spatial and temporal data together, the
joined dataset could be too large since each spatial variable will be
repeated at each time stamp for each site. Also, recordings of the site
ID from different data sources can be slightly different from each
other, causing a painful checking and cleaning of site IDs before the
join.

A cubble, in essence, wires both dimensions in the spatio-temporal data
into one object while provide two forms for manipulation the spatial and
temporal dimension separately.

When manipulating the spatial dimension it uses a nest form that:

@@

When manipulating the temporal dimension, it uses the long form that:

\begin{itemize}
\tightlist
\item
  each combination of group and timestamp occupies a row
\item
  time-related variables are displayed, and
\item
  group-related variables are not explicitly displayed but can be
  accessed through the \texttt{meta} attribute.
\end{itemize}

\hypertarget{create-a-cubble}{%
\section{Create a cubble}\label{create-a-cubble}}

The creation of a cubble requires the site identifier (\texttt{key}), as
well as the spatial (\texttt{coords}) and temporal (\texttt{index})
identifier. \texttt{climate\_flat} is already a tibble and it uses
\texttt{id} to identify each station, \texttt{date} as the time
identifier, and \texttt{c(long,\ lat)} as the spatial identifier. To
create a cubble for this data, use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{climate\_flat }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: id-wise: nested form
## # Key:    id [5]
## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
##   id            lat  long  elev name           wmo_id ts                
##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{verbatim}

Most of the time, spatio-temporal data doesn't come into this form and
analysts need to query the climate variables based on station metadata.
For this type of task, one can structure a metadata into a tibble and
use row-wise operator to query the climate variables into a nested list.
As an example here we demonstrate the workflow to find the 5 closest
stations to Melbourne. We first create a station data farme with the 5
target stations.

\begin{verbatim}
## # A tibble: 5 x 8
##   id            lat  long  elev name                 wmo_id  dist city     
##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
\end{verbatim}

We can query the climate information into a nested list named
\texttt{ts} for each station with the \texttt{rowwise()} operator. To
create a cubble, supply the same identifiers as with the first example.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sydmel\_climate }\OtherTok{\textless{}{-}}\NormalTok{ stations }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ts =} \FunctionTok{list}\NormalTok{(}\FunctionTok{meteo\_pull\_monitors}\NormalTok{(id, }
                                       \AttributeTok{date\_min =} \StringTok{"2020{-}01{-}01"}\NormalTok{, }
                                       \AttributeTok{date\_max =} \StringTok{"2020{-}12{-}31"}\NormalTok{,}
                                       \AttributeTok{var =} \FunctionTok{c}\NormalTok{(}\StringTok{"PRCP"}\NormalTok{, }\StringTok{"TMAX"}\NormalTok{, }\StringTok{"TMIN"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
                     \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{id))) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: id-wise: nested form
## # Key:    id [5]
## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
##   id            lat  long  elev name                 wmo_id  dist city   ts     
##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
\end{verbatim}

\hypertarget{cubble-operations}{%
\subsection{Cubble operations}\label{cubble-operations}}

\hypertarget{basics}{%
\subsubsection{Basics}\label{basics}}

\begin{itemize}
\tightlist
\item
  \texttt{stretch}: nest to long form
\item
  \texttt{tamp}: long to nest form
\item
  \texttt{migrate}: move selected spatial variables to the long form.
\item
  \texttt{add\_dscrb\_prct}: summary stats for missingness
\end{itemize}

dplyr compatibility:

\begin{itemize}
\tightlist
\item
  mutate, filter, summarise, select, arrange
\item
  group and ungroup: group\_by, ungroup
\item
  slice family
\end{itemize}

\hypertarget{combine-two-cubbles}{%
\subsubsection{Combine two cubbles}\label{combine-two-cubbles}}

\begin{itemize}
\tightlist
\item
  match river and weather gauges data
\item
  involve combining two cubbles
\item
  join operations combine the two together by appending more rows but
  what we really want is to bind rows.
\item
  bind rows also doesn't work since we want to bind only when there' s a
  matching????
\item
  introduce bind\_join
\end{itemize}

\hypertarget{hierarchical-structure-in-cubble}{%
\subsubsection{Hierarchical structure in
cubble}\label{hierarchical-structure-in-cubble}}

\begin{itemize}
\tightlist
\item
  hierarchical is common.
\item
  Given examples.
\item
  Essence: switch between different levels
\item
  introduce \texttt{switch\_key}
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{==} \DecValTok{2020}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{tamp}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: id-wise: nested form
## # Key:    id [85]
## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
##    id            lat  long  elev name                   wmo_id ts               
##    <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
##  1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
##  2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
##  3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
##  4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
##  5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
##  6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
##  7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
##  8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
##  9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
## 10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
## # ... with 75 more rows
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# join wmo\_id for each station}
\CommentTok{\# to\_join \textless{}{-} station \%\textgreater{}\% select(id, wmo\_id)}
\CommentTok{\# out \textless{}{-} climate\_small \%\textgreater{}\% }
\CommentTok{\#   left\_join(to\_join, by = c("station" = "id")) \%\textgreater{}\% }
\CommentTok{\#   filter(!is.na(wmo\_id))}
\CommentTok{\# out}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{migrate}\NormalTok{(id, lat, long)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: time-wise: long form
## # Key:    id [85]
## # Leaves: id [chr], lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
## #   [dbl], prcp_missing [dbl], tmax_missing [dbl], tmin_missing [dbl]
##    id          date        prcp  tmax  tmin   lat  long
##    <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
##  1 ASN00009021 2020-01-01     0   319   153 -31.9  116.
##  2 ASN00009021 2020-01-02     0   249   164 -31.9  116.
##  3 ASN00009021 2020-01-03     6   232   130 -31.9  116.
##  4 ASN00009021 2020-01-04     0   284   124 -31.9  116.
##  5 ASN00009021 2020-01-05     0   353   116 -31.9  116.
##  6 ASN00009021 2020-01-06     0   348   131 -31.9  116.
##  7 ASN00009021 2020-01-07     0   328   151 -31.9  116.
##  8 ASN00009021 2020-01-08     0   304   174 -31.9  116.
##  9 ASN00009021 2020-01-09     0   287   173 -31.9  116.
## 10 ASN00009021 2020-01-10     0   326   158 -31.9  116.
## # ... with 31,100 more rows
\end{verbatim}
